---
title: "Predicting Oscar Best Picture 2020"
author: "Joseph Shu, Jimena Rey, Georg Heimann, Ruby Zhou"
date: "1-2-2020"

---

### Load Packages 
```{r setup, include=FALSE}
library(xml2)
library(rvest)
library(stringr)
library(dplyr)
library(tidyverse)
library(openxlsx)
library(magrittr)
library(ggpubr)
library(tm)
```


```{r, message = F, warning = F}
#Specifying the url for desired website to be scrapped
url <- "https://en.wikipedia.org/wiki/Academy_Award_for_Best_Picture"

#Reading the html content and scrape data
webpage <- read_html(url)
table_html <- html_nodes(webpage, ".wikitable")
table <- html_table(table_html, fill = TRUE)

df <- table[[1]]

### bind rows to df

for ( i in 2:10){
  
  colnames(table[[i]]) <- c("Year", "Film", "Film Studio") 
  df <- rbind(df,table[[i]])
  
}
### delete NA rows
df <- df[ !is.na( df[,2] ) , ]

### winner is first row of each year


links <- tibble( 
  webpage %>%
    html_nodes(".wikitable") %>% html_nodes("i") %>% html_nodes("a") %>% 
    html_attr("href")
)
matchname <- tibble (
       webpage %>%
       html_nodes(".wikitable") %>% html_nodes("i") %>% html_nodes("a") %>% 
       html_text("href")
)

links_df <- data.frame(links,matchname)

colnames(links_df) <- c("Link","Filmcheck")

data <- data.frame(df, links_df)

## check if scraping worked
data$Film==data$Filmcheck
data <- data %>% select(-Filmcheck)

data$Link2 <- paste0("https://en.wikipedia.org",data$Link )

movielist <- data$Link2
IMDBlist <- list()

for (i in 1:length(movielist)){
  
   list <- read_html(movielist[i]) %>% html_nodes("ul li") %>% 
    html_nodes("a") %>% html_attr("href")

 IMDBlist[i]= list[grep('imdb', list)]
}

data2 <- data.frame(data, IMDBlink=unlist(IMDBlist))
### now we have all the imdb links to scrape the data
### unfortunately some mistakes occured that have to be changed manually

####changing mistakes
data2$IMDBlink<- as.character(data2$IMDBlink)
data2[data2$Link=="/wiki/Lincoln_(film)","IMDBlink"]<- "https://www.imdb.com/title/tt0443272/"
data2[data2$Link=="/wiki/Mr._Deeds_Goes_to_Town","IMDBlink"]<- "https://www.imdb.com/title/tt0027996/"
data2[data2$Link=="/wiki/Elizabeth_(film)","IMDBlink"]<- "https://www.imdb.com/title/tt0127536/"
data2[data2$Link=="/wiki/A_Room_with_a_View_(1985_film)","IMDBlink"]<- "https://www.imdb.com/title/tt0091867/"
data2[data2$Link=="/wiki/Going_My_Way","IMDBlink"]<- "https://www.imdb.com/title/tt0443272/"
data2[data2$Link=="/wiki/Three_Coins_in_the_Fountain_(film)","IMDBlink"]<- "https://www.imdb.com/title/tt0047580/"
data2[data2$Link=="/wiki/One_Hundred_Men_and_a_Girl","IMDBlink"]<- "https://www.imdb.com/title/tt0029347/"
data2[data2$Link=="/wiki/One_Hundred_Men_and_a_Girl","IMDBlink"]<- "https://www.imdb.com/title/tt0029347/"
data2[data2$Link=="/wiki/Casablanca_(film)","IMDBlink"]<- "https://www.imdb.com/title/tt0034583/"
```


```{r}
###scrape rating and time
###takes around 15 min

IMDBsites <- data2$IMDBlink
IMDBrating <- list()
IMDBtime <- list()

i=1
while (i <= length(IMDBsites)){
  x= print(read_html(IMDBsites[i]) %>% html_nodes(".ratingValue") %>% html_text() %>% gsub("\n","",.))
  if ( identical(x, character(0)) ) {x <- NA}
  IMDBrating[i]<- x
  
  y= print((read_html(IMDBsites[i]) %>% html_nodes("time") %>% html_text())[1] %>% gsub("\n ","",.))
  if ( identical(x, character(0)) ) {x <- NA}
  IMDBtime[i]<- y
  
  i=i+1
}
data2 <- data.frame(data2, IMDBrating=unlist(IMDBrating), Time=unlist(IMDBtime))
data2$Time <- data2$Time %>% str_replace_all(pattern=" ",replacement = "")
```

```{r}
### scrape Genre

IMDBGenre1 <-list()
IMDBGenre2 <-list()
IMDBGenre3 <- list()
IMDBGenre4 <-list()
i=1
while (i <= length(IMDBsites)){
  x= read_html(IMDBsites[i]) %>% html_nodes(".subtext a") %>% html_text()
  if ( identical(x, character(0)) ) {x <- NA}
  IMDBGenre1[i]=print(x[1])
  IMDBGenre2[i]=print(x[2])
  IMDBGenre3[i]=print(x[3])
  IMDBGenre4[i]=print(x[4])
i=i+1
}
Genre.data <- data.frame(G1=unlist(IMDBGenre1),G2=unlist(IMDBGenre2),G3=unlist(IMDBGenre3),G4=unlist(IMDBGenre4))
Genre.data$Release <- as.character(NA)
Genre.data$G2 <- as.character(Genre.data$G2)
Genre.data$G3 <- as.character(Genre.data$G3)
Genre.data$G4 <- as.character(Genre.data$G4)
Genre.data$G1 <- as.character(Genre.data$G1)

Genre.data[which(str_detect(Genre.data$G2,pattern="[1-9]")==TRUE),"Release"] <-
  Genre.data[which(str_detect(Genre.data$G2,pattern="[1-9]")==TRUE),"G2"]
Genre.data[which(str_detect(Genre.data$G2,pattern="[1-9]")==TRUE),"G2"] <- NA

Genre.data[which(str_detect(Genre.data$G3,pattern="[1-9]")==TRUE),"Release"] <-
  Genre.data[which(str_detect(Genre.data$G3,pattern="[1-9]")==TRUE),"G3"]
Genre.data[which(str_detect(Genre.data$G3,pattern="[1-9]")==TRUE),"G3"] <- NA

Genre.data[which(str_detect(Genre.data$G4,pattern="[1-9]")==TRUE),"Release"] <-
  Genre.data[which(str_detect(Genre.data$G4,pattern="[1-9]")==TRUE),"G4"]
Genre.data[which(str_detect(Genre.data$G4,pattern="[1-9]")==TRUE),"G4"] <- NA

data2 <- data.frame(data2, Genre1=Genre.data$G1, Genre2=Genre.data$G2, Genre3=Genre.data$G3)
```

```{r}
### Winner is first row of each year
winnerlist <- list()
b <- c( as.character(data2$Year))
for( a in 1:562){winnerlist[a]=(b[a]!=data2$Year[a+1])}
winnerlist <- c(TRUE, unlist(winnerlist)) %>% as.numeric()
data2$winner <- winnerlist
data2[555,"winner"]<-0
```
```{r}
### scrape number of votes and budget
IMDBvotes <- list()

i=1
while (i <= length(IMDBsites)){
  x= (read_html(IMDBsites[i]) %>% html_nodes(".small")%>% html_text("span"))[2]
  if ( identical(x, character(0)) ) {x <- NA}
  IMDBvotes[i]<- x
  
  i=i+1
}

data2$Votes <- unlist(IMDBvotes)
data2[8,"Votes"]<- NA


#### budget
IMDBbudget <- list()

data2$Votes <- IMDBvotes
i=1
while (i <= length(IMDBsites)){
  x= (read_html(IMDBsites[i]) %>% html_nodes("#titleDetails")%>%  html_text())
  if ( identical(x, character(0)) ) {x <- NA}
  x1 <- sub(".*Budget:", "", x)
  x1 <- sub("\\n.*","",x1)
  IMDBbudget[i]<- print(x1)
  
  i=i+1
}

data2$Budget <- unlist(IMDBbudget)
```

```{r}
### rotten tomatoes link
rottenlist <- list()

for (i in 1:length(movielist)){
  
  x <- read_html(movielist[i]) %>% html_nodes("ul li") %>% 
    html_nodes("a") %>% html_attr("href")
  x <-print(x[grep('rotten', x)])
  if ( identical(x, character(0)) ) {x <- NA}
  rottenlist[i]= x
}

data2$rottenlink <- unlist(rottenlist)



###rotten tomatoes rating
rottenrating1 <-list()
rottenrating2 <-list()
i=1
```

```{r}
### This chunk has to be executed until no error occurs anymore

while (i <= length(rottenlist)){
  
  if (is.na(rottenlist[i])){
    rottenrating1[i]<- NA
    rottenrating2[i]<- NA
  }else
  {
    
    x= (read_html(rottenlist[[i]]) %>% html_nodes("#tomato_meter_link .mop-ratings-wrap__percentage")%>%  html_text())
    if ( identical(x, character(0)) ) {x <- NA}
    x <- gsub("[^0-9.]", "",  x)
    rottenrating1[i]<- print(x)
    
    x2= (read_html(rottenlist[[i]]) %>% html_nodes(".audience-score .mop-ratings-wrap__percentage")%>%  html_text())
    if ( identical(x, character(0)) ) {x2 <- NA}
    x2 <- gsub("[^0-9.]", "",  x2)
    rottenrating2[i]<- print(x2)
    
  }
  i=i+1
}
### 4 errors come... do this 3 steps and start loop again
rottenrating1[i]<- NA
rottenrating2[i]<- NA
i=i+1

```

```{r}
data2$tomatometer <- unlist(rottenrating1)
data2$audiencescore <- unlist(rottenrating2)

### write xlsx not clean
#write.xlsx(data2, "data.xlsx")
```

### Cleaning Data
```{r}
# year
data2$Year <-sub("\\(.*","",data2$Year)
data2$Year <- sub("\\/.*","",data2$Year)

# imdb rating
data2$IMDBrating <-sub("\\/.*","",data2$IMDBrating)

#time
data2$Time <- sapply(str_extract_all(data2$Time, "\\d+"), function(x) {
  x1 <- as.numeric(x)
  if(length(x1)>1) x1[1]*60 + x1[2] else x1*60 })
### Casablanca is wrong

# Votes
data2$Votes <- as.numeric(gsub(",", "", data2$Votes))

data2$Year <- data2$Year %>% as.numeric
data2$IMDBrating <- as.numeric(data2$IMDBrating)
data2$tomatometer <- as.numeric(data2$tomatometer)
data2$audiencescore <- as.numeric(data2$audiencescore)


nominee <- data2
```


```{r}
# Budget
nominee$Currency <-nominee$Budget
nominee$Currency <- removeNumbers(nominee$Currency)
nominee$Currency <- gsub(",","",nominee$Currency)
nominee$Budget <- parse_number(nominee$Budget)
#Convert GBP and EUR to USD - convertion rates gathered from various websites
Y1933 <- nominee %>% 
  filter(Currency == "GBP", Year == 1933) %>% 
  mutate(Budget=Budget/0.24)
Y1938 <- nominee %>% 
  filter(Currency == "GBP", Year == 1938) %>% 
  mutate(Budget=Budget/0.2)
Y1942_48 <- nominee %>% 
  filter(Currency == "GBP", Year %in% c(1942, 1943, 1946, 1947,1948)) %>% 
  mutate(Budget=Budget/0.2525)
Y1959 <- nominee %>% 
  filter(Currency == "GBP", Year == 1959) %>% 
  mutate(Budget=Budget/0.3508)
Y1965 <- nominee %>% 
  filter(Currency == "GBP", Year == 1965) %>% 
  mutate(Budget=Budget/0.3517)
Y1983_87 <- nominee %>% 
  filter(Currency == "GBP", Year %in% c(1983,1987)) %>% 
  mutate(Budget=Budget/0.4303)
Y1989 <- nominee %>% 
  filter(Currency == "GBP", Year == 1989) %>% 
  mutate(Budget=Budget/0.6392)
Y1992 <- nominee %>% 
  filter(Currency == "GBP", Year == 1992) %>% 
  mutate(Budget=Budget/0.6015)
Y2006 <- nominee %>% 
  filter(Currency == "GBP", Year == 2006) %>% 
  mutate(Budget=Budget*1.84)
Y2017 <- nominee %>%
  filter(Currency == "EUR") %>%
  mutate(Budget=Budget*1.13)

GbpEurYears <- rbind(Y1933,Y1938,Y1942_48,Y1959,Y1965,Y1983_87,Y1989,Y1992,Y2006, Y2017)
GbpEurYears$Currency <- sub("GBP", "$", GbpEurYears$Currency)
GbpEurYears$Currency <- sub("EUR", "$", GbpEurYears$Currency)

#Merging GBPYears to nominee and deleting duplicates with GBP and EUR values. Since now all budget values are in USD, will delete Currency column
nominee <- rbind(nominee,GbpEurYears)
nominee <- nominee[!(nominee$Currency %in% c("GBP","EUR")),]
nominee <- select(nominee, -"Currency")
```


```{r}
# Read html table
url <- "https://inflationdata.com/Inflation/Consumer_Price_Index/HistoricalCPI.aspx?reloaded=true"
webpage <- read_html(url)
USinflation <- html_table(webpage, fill = TRUE)
USinflation <- USinflation[[1]]
USinflation <- USinflation %>% select(Year , Ave.)
# Rename column for clarity
names(USinflation)[names(USinflation) == "Ave."] <- "InflationR"
# View resulting dataframe
head(USinflation)
```

```{r}
#Adjust Budget to today's values
nominee <- merge(nominee, USinflation, by = "Year" )
AdjustedBudget <- with(nominee, Budget*(255.657/InflationR))
nominee <- cbind(nominee, AdjustedBudget)
nominee <- select(nominee, -c("InflationR","Budget"))
```

```{r}
# Dealing with missing values
NAcols <- colnames(nominee)[colSums(is.na(nominee)) > 0]
sort(colSums(sapply(nominee[NAcols], is.na)))
```

```{r}
#Votes
nominee$Votes[is.na(nominee$Votes)] <- mean(nominee$Votes[!is.na(nominee$Votes)])
nominee$IMDBrating[is.na(nominee$IMDBrating)] <- mean(nominee$IMDBrating[!is.na(nominee$IMDBrating)])

```

```{r}
# AdjustedBudget - ratio imputation
model3 <- lm(AdjustedBudget ~ IMDBrating + Votes, data = nominee)
summary(model3)
NAadjustedbudget <- is.na(nominee$AdjustedBudget)
nominee$AdjustedBudget[NAadjustedbudget] <- predict(model3, newdata = nominee[NAadjustedbudget, ])
```

```{r}
# tomatometer - ratio imputation
model4 <- lm(tomatometer ~ IMDBrating + Time + Votes, data = nominee)
summary(model4)
NAtomatometer <- is.na(nominee$tomatometer)
nominee$tomatometer[NAtomatometer] <- predict(model4, newdata = nominee[NAtomatometer, ])

nominee <- select(nominee, -rottenlink)
```

```{r}
#AudienceScore
model5 <-lm(audiencescore ~ tomatometer + IMDBrating + Votes, data = nominee)
summary(model5)
NAaudience <- is.na(nominee$audiencescore)
nominee$audiencescore[NAaudience] <- predict(model5, newdata = nominee[NAaudience, ])

### write.xlsx(nominee,"dataclean 01.02.20.xlsx")
nominee <- read_excel("dataclean 01.02.20.xlsx")
```

### Data Visulisation
```{r}
### Genre
nominee$Year <- as.numeric(nominee$Year)

G1 <-nominee %>%  group_by(Genre1) %>% summarise(n=n())
G2 <-nominee %>%  group_by(Genre2) %>% summarise(n=n()) %>% rename(Genre1="Genre2")
G3 <-nominee %>%  group_by(Genre3) %>% summarise(n=n()) %>% rename(Genre1="Genre3")
G <- rbind(G1,G2,G3) %>% group_by(Genre1) %>% summarise(sum=sum(n))

## junst winner genre
datawinner <- nominee %>% filter(winner==1)
G1winner <-datawinner %>%  group_by(Genre1) %>% summarise(n=n())
G2winner <-datawinner %>%  group_by(Genre2) %>% summarise(n=n()) %>% rename(Genre1="Genre2")
G3winner <-datawinner %>%  group_by(Genre3) %>% summarise(n=n()) %>% rename(Genre1="Genre3")
Gwinner <- rbind(G1winner,G2winner,G3winner) %>% group_by(Genre1) %>% summarise(sum=sum(n))

G <- data.frame(Genre =c(nominee$Genre1,nominee$Genre2,nominee$Genre3),Winner=as.factor(c(nominee$winner,nominee$winner,nominee$winner)),Year=c(nominee$Year,nominee$Year,nominee$Year)) %>% na.omit()

###dist genre
ggplot(G, aes(reorder(Genre,Genre,function(x)-length(x)),fill=Winner)) + 
  geom_bar()+labs(title="Nominations by Genre",
        x ="Genre", y = "n")+
  theme(axis.text.x = element_text(face="bold", color="#993333", size=14),
        axis.text.y = element_text(face="bold", color="#993333", size=14))+coord_flip()

```

```{r}
### make new categories that make sense
nominee$Crime <- 0
nominee[which(nominee$Genre1=="Crime"),"Crime"]<-1
nominee[which(nominee$Genre1=="Thriller"),"Crime"]<-1
nominee[which(nominee$Genre1=="Film-Noir"),"Crime"]<-1
nominee[which(nominee$Genre2=="Crime"),"Crime"]<-1
nominee[which(nominee$Genre2=="Thriller"),"Crime"]<-1
nominee[which(nominee$Genre2=="Film-Noir"),"Crime"]<-1
nominee[which(nominee$Genre3=="Crime"),"Crime"]<-1
nominee[which(nominee$Genre3=="Thriller"),"Crime"]<-1
nominee[which(nominee$Genre3=="Film-Noir"),"Crime"]<-1

nominee$Romance <- 0
nominee[which(nominee$Genre1=="Romance"),"Romance"]<-1
nominee[which(nominee$Genre2=="Romance"),"Romance"]<-1
nominee[which(nominee$Genre3=="Romance"),"Romance"]<-1

nominee$Action <- 0
nominee[which(nominee$Genre1=="Action"),"Action"]<-1
nominee[which(nominee$Genre1=="Adventure"),"Action"]<-1
nominee[which(nominee$Genre1=="War"),"Action"]<-1
nominee[which(nominee$Genre1=="Sci-Fi"),"Action"]<-1
nominee[which(nominee$Genre2=="Action"),"Action"]<-1
nominee[which(nominee$Genre2=="Adventure"),"Action"]<-1
nominee[which(nominee$Genre2=="War"),"Action"]<-1
nominee[which(nominee$Genre2=="Sci-Fi"),"Action"]<-1
nominee[which(nominee$Genre3=="Action"),"Action"]<-1
nominee[which(nominee$Genre3=="Adventure"),"Action"]<-1
nominee[which(nominee$Genre3=="War"),"Action"]<-1
nominee[which(nominee$Genre3=="Sci-Fi"),"Action"]<-1

nominee$History <- 0
nominee[which(nominee$Genre1=="Biography"),"History"]<-1
nominee[which(nominee$Genre1=="Documentary"),"History"]<-1
nominee[which(nominee$Genre1=="History"),"History"]<-1
nominee[which(nominee$Genre1=="Music"),"History"]<-1
nominee[which(nominee$Genre2=="Biography"),"History"]<-1
nominee[which(nominee$Genre2=="Documentary"),"History"]<-1
nominee[which(nominee$Genre2=="History"),"History"]<-1
nominee[which(nominee$Genre2=="Music"),"History"]<-1
nominee[which(nominee$Genre3=="Biography"),"History"]<-1
nominee[which(nominee$Genre3=="Documentary"),"History"]<-1
nominee[which(nominee$Genre3=="History"),"History"]<-1
nominee[which(nominee$Genre3=="Music"),"History"]<-1

nominee$Comedy <- 0
nominee[which(nominee$Genre1=="Comedy"),"Comedy"]<-1
nominee[which(nominee$Genre2=="Comedy"),"Comedy"]<-1
nominee[which(nominee$Genre3=="Comedy"),"Comedy"]<-1

nominee <- nominee %>% select(-c("Genre1","Genre2","Genre3","Film.Studio"))
```

```{r}
hist(nominee$Time)
hist(nominee$IMDBrating)
hist(nominee$Votes) ### votes are screw
hist(nominee$AdjustedBudget)
### one outlier
boxplot.stats(nominee$AdjustedBudget) ###replace outlier with extreme of the upper ‘hinge’
nominee[nominee$Film=="Going My Way","AdjustedBudget"]=nominee[(nominee$Film=="Cleopatra"&nominee$Year==1963),"AdjustedBudget"]

### Give the movie also the second highest value to keep it expensive 
```

We start by looking at the summary statistics to give us an idea of what the dataset looks like
```{r}
#View summary statistics for all variables
summary(nominee)
```

We then plot a correlation matrix of the numeric variables
```{r, message = F, warning =F}
# View correlation between numeric variables of our dataset
ggpairs(nominee[ , c("winner","IMDBrating","Time","Votes","tomatometer","audiencescore","AdjustedBudget")])
```
From the matrix above we can see the only two variables that have a high correlation are audience score and IMDB rating, this is not too surprising since both these ratings are determined by normal viewers. Another variable that scores movies is tomatometer, message = F, warning =F which refers to movie critics. It is interesting to see this rating is not too closely correlated with either audience score nor the IMDB rating.

```{r, message = F, warning =F }
#Historical progression of movie budgets
ggplot(nominee) +
  geom_point(aes(Year, AdjustedBudget, colour = as.factor(winner))) +
  geom_smooth(aes(Year, AdjustedBudget))+
  theme_minimal() +
  labs(title = "Did movies get more expensive over time?", y = "Budget", colour = "Winner")
```
There is a trend for higher and higher budgets as years pass by. We also notice in the 2000s those movies with very large budgets do not seem to be more likely to win oscars for the most part.

```{r, message = F}
# Historical progression of movie length in minutes
ggplot(nominee) +
  geom_point(aes(Year, Time, colour = as.factor(winner))) +
  geom_smooth(aes(Year, Time))+
  theme_minimal() +
  labs(title = "Did movies get  longer over time?", y = "Movie Time in minutes", colour = "Winner")
```
As evident from the trend line, the movies are getting longer in general, although they seem to have plateaued after the 1950s. However, we observe more outliers between 1928 and 1975 than between 1975 and 2019. We also see that ever since around 2020 those movies which are much longer than the average seem to rarely win the Oscar for best picture.

```{r, message = F, warning =F}
#Relationship between movie length and budget
ggplot(nominee, aes(Time, AdjustedBudget)) +
  geom_point() +
  geom_smooth()+
  theme_minimal() +
  labs(title = "Are longer movies more expensive?", x="Movie Time in minutes", y="Budget")
```
Although the relationship is positive it is not very strong, this indicates that longer movies were not necessarily more expensive to produce. 

```{r, message = F, warning =F}
# Realtionship between IMDB votes and award-winning movies
ggplot(nominee, aes(x=winner, y=Votes, color = as.factor(winner))) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8, nominee.size=4) +
  theme_minimal() + 
 labs(title = "Did the winning movies get a more votes on IMDB?", x="Did the movie win?", colour = "Winner")

```
On average winning moves have higher number of votes on IMDB and less nominees than lose than lost. We need to bare in mind that this voting numbers are up to the date the data was scrapped. So it is possible after being nominated more people watched the movie and voted, thus skewing our results. This graph made us wonder about the relationship between votes and IMDB ratings, so we will analyse this next.

```{r, message = F, warning =F}
#Relationship between IMDB rating and votes
ggplot(nominee) +
  geom_point(aes(IMDBrating, Votes, colour = as.factor(winner))) +
  geom_smooth(aes(IMDBrating, Votes))+
  theme_minimal() +
  labs(title = "Do highest rated movies have more votes?", colour = "Winner")
```
The graph above shows there is definitely a positive relationship between highest rated movies and votes. We can also see an nominee at the top right, which had a very high rating and many votes, yet it did not win an Oscar. It is noteworthy that where the bulk of the movies are - rated between 7 and 8 - the regression line is almost horizonal indicating a neutral relationship.

```{r, message = F, warning = F}
#Relationship between movie length and IMDB rating
ggplot(nominee) + 
  geom_jitter(aes(IMDBrating, Time, colour = as.factor(winner))) +
  geom_smooth(aes(IMDBrating, Time, colour = winner)) + 
  theme_minimal() +
  labs(title = "Do longer movies get better IMDB ratings?", y = "Movie Lenght", colour = "Winner")
```
The trend line indicates that longer movies do receive better IMDB ratings. However, for where the bulk of the movies are - rated between 7 and 8 - the trend line is almost horizontal meaning the realtionship is not so positive for this average ratings. On the top right corner of the graph is also interesting to see that very long movies with very high ratings also won the Oscars. 

```{r, message = F, warning =F}
#Relationship winning movies and Rotten Tomatoes audience score
ggplot(nominee, aes(x=as.factor(winner), y=audiencescore, colour = as.factor(winner))) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8,nominee.size=4) +
 labs(title = "Did the winning movies get a higher Audience Score?", x = "Did the movie win?", colour = "Winner")

```
Although the average audience score for winning movies is higher, message = F, warning =F the difference is not that large. The outliers of winning movies are also more spread out. Seems like audience scores are not so good at predicting winning movies on their own.

```{r, message = F, warning =F }
#Genre groupings and audience scores
#Romance
ggplot(nominee, aes(x=Romance, y=audiencescore, color = Romance, group = Romance)) + 
  geom_boxplot(nominee.colour="blue", nominee.shape=8,nominee.size=4) +
  theme_minimal() + 
  theme(legend.position = "none") +
 labs(title = "Did Romance movies get a higher Audience Score?")

#Action
ggplot(nominee, aes(x=Action, y=audiencescore, color = Action, group = Action)) + 
  geom_boxplot(nominee.colour="blue", nominee.shape=8,nominee.size=4) +
  theme_minimal() + 
  theme(legend.position = "none") +
   labs(title = "Did Action movie get a higher Audience Score?")

#Crime
ggplot(nominee, aes(x=Crime, y=audiencescore, color = Crime, group = Crime)) + 
  geom_boxplot(nominee.colour="blue", nominee.shape=8, nominee.size=4)+
  theme_minimal() + 
  theme(legend.position = "none") +
  labs(title = "Did Crime movies get a higher Audience Score?")

#History
ggplot(nominee, aes(x=History, y=audiencescore, color = History, group = History)) + 
  geom_boxplot(nominee.colour="blue", nominee.shape=8,nominee.size=4) +
  theme_minimal() + 
  theme(legend.position = "none") +
    labs(title = "Did History movies get a higher Audience Score?")

#Comedy
ggplot(nominee, aes(x=Comedy, y=audiencescore, color = Comedy, group = Comedy)) + 
  geom_boxplot(nominee.colour="blue", nominee.shape=8,nominee.size=4) +
  theme_minimal() + 
  theme(legend.position = "none") +
    labs(title = "Did Comedy movies get a higher Audience Score?")

```
The first graph seems to indicate the audiences and the Academy Awards have different preferences when it comes to romantic movies, with the audience scoring higher those movies which did not end up winning an award. In regards to action movies, it seems like the audience enjoys them regardless of whether they ended up winning an award or not. Although we also see there are significantly more outliers for the action movies which did not win. For crime movies, the audiences seems to have a good feel of which movies will win, since these have higher average ratings and less deviation. For history the scores are pretty similar as well as the spread for both winning and losing movies, althouh losing ones have pretty distant outliers. Finally, for comedy movies both categories have similar average scores but winning movies have a larger deviation and no outliers. 
Since we saw in the correlation matrix that IMDBratings and audiencescore are highly correlated, the trends observed for genres in IMDB ratings are the same for the most part.


```{r, message=F, warning=F}
#Genre groupings and critics tomato meter
#Romance
ggplot(nominee, aes(x=Romance, y=tomatometer,color = Romance, group = Romance)) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8, nominee.size=4) +
    theme_minimal() + 
  theme(legend.position = "none") + 
  labs(title = "Did Romance movies get a higher tomatometer score?")

#Action
ggplot(nominee, aes(x=Action, y=tomatometer,  color = Action, group = Action)) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8, nominee.size=4) +
    theme_minimal() + 
  theme(legend.position = "none") +
  labs(title = "Did Action movies get a higher tomatometer score?")

#Crime  
ggplot(nominee, aes(x=Crime, y=tomatometer, color = Crime, group = Crime)) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8,nomminee.size=4) + 
    theme_minimal() + 
  theme(legend.position = "none") +
  labs(title = "Did Crime movies get a higher tomatometer score?")

#History
ggplot(nominee, aes(x=History, y=tomatometer, color = History, group = History)) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8,nominee.size=4) + 
    theme_minimal() + 
  theme(legend.position = "none") +
  labs(title = "Did History movies get a higher tomatometer score?")

#Comedy
ggplot(nominee, aes(x=Comedy, y=tomatometer, color = Comedy, group = Comedy)) + 
  geom_boxplot(nominee.colour="red", nominee.shape=8,nominee.size=4) +
    theme_minimal() + 
  theme(legend.position = "none") +
  labs(title = "Did Comedy movies get a higher tomatometer score?")


```
We are now looking at the third type of rating in our data set, that is the one by movie critics in rotten tomatoes, called the tomato meter. For romance movies we can see they score higher those which did not win and their scores have less deviation although far more outliers. For action movies the average scores are pretty similar but winning movies have a wider deviation. Regarding crime movies critics seem to be in agreement with the academy as to what makes a good crime movie, with a higher average score and much lower deviation with barely any outliers. History movies are interesting in the sense that although they have similar average secores they have a wider interquartile range and a lower minimum score. Finally, comedy movies have the same average scores, losing movies have a wider interquartile range as well as a lower minimum. For all five genre categories movies which did not win the Best Picture award have many more outliers than those which did win.


```{r, message = F, warning =F}
ggplot(nominee) +
  geom_point(aes(IMDBrating, tomatometer, colour = as.factor(winner))) +
  geom_smooth(aes(IMDBrating, tomatometer)) + 
  theme_minimal() + 
  labs(colour = "Winner", title = "Relationship between IMDB ratings and the Tomatometer")

```
In this graph, we can observe that for all the past winners, IMDBrating and tomatometer have a positive relationship, That said, winners who got higher ratings will usually get higher tomato meters.

```{r}
#Predictive power of IMDB rating, TomatoMeter and Audience Scores
summary(lm(winner ~ IMDBrating + tomatometer + audiencescore, data = nominee))
```
In the model above we compared our three rating measures and interestingly we can see that tomatometer actually predicts worse than audiencescore and IMDB ratings, even though it is from professional critics. The best predictor is IMDBratings followed by audience score, while the tomato meter comes in last. Even then all three measures together can only explain 4% of the variability of the winning movies. 

```{r}
#Models
nominee <- readRDS("nominee.Rds")

data <- filter(nominee,Year!=2019) %>% select(-Film)
prediction_data<- filter(nominee,Year==2019)

# define train and test data
set.seed(2)
train.n <- sample(1:length(data$Year),300)
train_df <- data[train.n,]
test_df <- data[-train.n,]

Win_ <- test_df$winner

#Logit
mylogit <- glm(winner ~., data = train_df, family = binomial)
summary(mylogit)
predicted = predict(mylogit, newdata=test_df, type="response")
#cv_error <- cv.glm(train_df,predicted)
summary(predicted)

cv.glm(train_df, mylogit, K=3)$delta[1]

logit_pred <- predicted
logit_pred[logit_pred>0.35]=1
logit_pred[logit_pred<0.35]=0

table(logit_pred,Win_)

predict(mylogit, newdata=prediction_data, type="response")
### The joker would win after logit

### probit
myprobit <- glm(winner ~., data = train_df, family = binomial(link = "probit"))
summary(myprobit)
predicted2 = predict(myprobit, newdata=test_df, type="response")
#cv_error <- cv.glm(train_df,predicted)
summary(predicted2)
cv.glm(train_df, myprobit, K=3)$delta[1]

probit_pred <- predicted2
probit_pred[probit_pred>0.35]=1
probit_pred[probit_pred<0.35]=0

table(probit_pred,Win_)

predict(myprobit, newdata=prediction_data, type="response")
### also number 5 Joker wins


############################################ tree based

tree_nominee <- tree(as.factor(winner)~.,data=data, subset = train.n)

### cross validation for optimal pruning
cv_nominee <- cv.tree(tree_nominee,FUN=prune.misclass)
best_level <- which.min(cv_nominee$dev)

prune_nominee <- prune.misclass(tree_nominee,best=best_level)

tree_pred=predict(prune_nominee,test_df,type="class")
table(tree_pred,Win_)

predict(prune_nominee, newdata=prediction_data, type="class")

############################################random forest

randomforest_nominee <- randomForest(as.factor(winner)~.,data=data, subset = train.n, mtry=13,ntree=11)

randomforest_pred=predict(randomforest_nominee,test_df,type="class")
table(randomforest_pred,Win_)

predict(randomforest_nominee, newdata=prediction_data, type="class")
### also 5 wins but winner changes easily with changing the ntree
#################################################
```

